Limitations on the Two-Infall Scenario in the Era of Large Stellar Age Catalogs

### 2025-03-26 src/tex/ms.tex
Updates from Overleaf

\begin{figure*}
    \centering
    \includegraphics{src/tex/figures/ofe_df_comparison.pdf}
    \caption{Stellar [O/Fe] distributions produced by select multi-zone models (a--d) and as observed by APOGEE (e). Each row presents stellar distributions within a range of absolute midplane distance $|z|$ reported on the far right. Within each panel, the distributions are color-coded according to the bin in galactocentric radius $R_{\rm gal}$ from which they are drawn. Each distribution is normalized so that the area under the curve is 1, and the vertical scale is consistent across each row. The median APOGEE abundance uncertainties are forward-modeled onto the model outputs (see Table \ref{tab:uncertainties}). For visual clarity, each distribution is smoothed with a box-car of width 0.05 dex. Model (a) assumes the fiducial parameters and the \yZ{1} yield set, whereas models (b--d) assume the \yZ{2} yield set. Model (c) also adopts a power-law SN Ia DTD, while in model (d) the infalling gas has metallicity $\mathOH_{\rm CGM}=\mathFeH_{\rm CGM}=-0.7$. {\bf Key takeaway:} The two-infall model produces an intermediate-[O/Fe] peak that is especially prominent when assuming higher yields and outflows. This effect can be mitigated by pre-enriched infall.}
    \label{fig:ofe-df}
\end{figure*}

@@

@@

The dilution effect discussed in Section \ref{sec:yields} is clearly seen in the multi-zone model results. Figure \ref{fig:abundance-evolution} shows stellar age--abundance relations for four model outputs. Column (a) plots the output of our fiducial model, with $y/Z_\odot=1$, $\sigma_{\rm RM8}=2.7\,{\rm kpc}$, and $Z_{\rm CGM}=0$. The fiducial model shows two major discrepancies with the \citet{leung_variational_2023} age--abundance relation: a major dilution of $\sim0.5$ dex at a lookback time of $\sim9\,{\rm Gyr}$ near where the data show a maximum in [O/H], and non-zero abundance evolution at late times where the data show very little abundance evolution. The evolution of [Fe/H] is similar, but the approach to Solar abundance is slower because of the additional delay imposed on Fe production from SNe Ia. A higher yield set (column b) mitigates both of these issues by shortening the time it takes the ISM metallicity to rebound post-$t_{\rm max}$, producing a much flatter abundance curve at late times. However, model (b) produces a poorer fit to the age--[O/Fe] relation; the decline in [O/Fe] over the thin disk epoch is much steeper than the data.

The observed rise in the median abundance of stars with ages of $\sim4-8\,{\rm Gyr}$ is thought to be due to radial migration, as those stars are more likely to have migrated from the dense inner metal-rich regions of the Galaxy during that time. Although our fiducial model does include a prescription for radial migration, the majority of stars in that age range in column (a) have sub-Solar abundances. Column (c) of Figure \ref{fig:abundance-evolution} presents a model with $y/Z_\odot=1$ but with a stronger migration prescription of $\sigma_{\rm RM8}=5\,{\rm kpc}$. Here, the stars which make up the present-day Solar neighborhood are drawn from a wider range of birth $R_{\rm gal}$, producing a broader abundance distribution for any given age. However, even though this prescription is much stronger than the estimates of, e.g., \citet{frankel_measuring_2018}, the model cannot reproduce the observed age--abundance trends in [O/H] and [Fe/H].

Finally, as in Section \ref{sec:abundance-distributions}, we investigate a model where the gas infall is pre-enriched to $\mathOH=\mathFeH=-0.7$. As shown in column (d) of Figure \ref{fig:abundance-evolution}, this mitigates but does not completely solve the two discrepancies. The dilution effect of the second infall is reduced to the $\sim0.3$-dex level as the gas which replenishes the Galaxy's reservoir is no longer pristine; however, the width of the stellar abundance distribution at any given age is also reduced, because the enriched gas accretion imposes a lower limit on the metallicity of the outermost regions, from which the stars of the low-metallicity tail in the Solar neighborhood are drawn. The late-time gas abundance evolution is similar to the fiducial model, but it ends at slightly super-Solar metallicity --- an effect which can be compensated by a slightly increased value of $\eta$. Overall, no modification to the $y/Z_\odot=1$ model is able to overcome both the dilution and late-time evolution issues.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{src/tex/figures/ofe_distribution_yields.pdf}
    \caption{Caption}
    \label{fig:ofe-distribution-yields}
\end{figure}

\begin{figure*}
    \centering
    \includegraphics{src/tex/figures/ofe_distribution_params.pdf}
    \caption{Stellar [O/Fe] distributions produced by select multi-zone models (a--d) and as observed by APOGEE (e). Each row presents stellar distributions within a range of absolute midplane distance $|z|$ reported on the far right. Within each panel, the distributions are color-coded according to the bin in galactocentric radius $R_{\rm gal}$ from which they are drawn. Each distribution is normalized so that the area under the curve is 1, and the vertical scale is consistent across each row. The median APOGEE abundance uncertainties are forward-modeled onto the model outputs (see Table \ref{tab:uncertainties}). For visual clarity, each distribution is smoothed with a box-car of width 0.05 dex. Model (a) assumes the fiducial parameters and the \yZ{1} yield set, whereas models (b--d) assume the \yZ{2} yield set. Model (c) also adopts a power-law SN Ia DTD, while in model (d) the infalling gas has metallicity $\mathOH_{\rm CGM}=\mathFeH_{\rm CGM}=-0.7$. {\bf Key takeaway:} The two-infall model produces an intermediate-[O/Fe] peak that is especially prominent when assuming higher yields and outflows. This effect can be mitigated by pre-enriched infall.}
    \label{fig:ofe-df}
\end{figure*}

@@

The dilution effect discussed in Section \ref{sec:yields} is clearly seen in the multi-zone model results. Figure \ref{fig:abundance-evolution-yields} shows stellar age--abundance relations produced by models with \yZ{1} and \yZ{2} with fiducial parameters (Table \ref{tab:multizone-parameters}). The \yZ{1} model (column a) shows two clear discrepancies with the \citet{leung_variational_2023} age--abundance relation: a major dilution of $\sim0.5$ dex at a lookback time of $\sim9\,{\rm Gyr}$ near where the data show a maximum in [O/H], and non-zero abundance evolution at late times where the data show very little abundance evolution. The evolution of [Fe/H] is similar, but the approach to Solar abundance is slower because of the additional delay imposed on Fe production from SNe Ia. The \yZ{2} yield set (column b) mitigates both of these issues by shortening the time it takes the ISM metallicity to rebound post-$t_{\rm max}$, producing a much flatter abundance curve at late times. However, model (b) produces a poorer fit to the age--[O/Fe] relation: the decline in [O/Fe] over the thin disk epoch is much steeper than the data.

@@

Figure \ref{fig:abundance-evolution-params} shows the effect of varying several parameters on the \yZ{1} model.

The observed rise in the median abundance of stars with ages of $\sim4-8\,{\rm Gyr}$ is thought to be due to radial migration, as those stars are more likely to have migrated from the dense inner metal-rich regions of the Galaxy during that time. Although our fiducial model does include a prescription for radial migration, the majority of stars in that age range in Figure \ref{fig:abundance-evolution-yields} still have sub-Solar abundances. Column (b) of Figure \ref{fig:abundance-evolution-params} presents a model with $y/Z_\odot=1$ but with a stronger migration prescription of $\sigma_{\rm RM8}=5\,{\rm kpc}$. Here, the stars which make up the present-day Solar neighborhood are drawn from a wider range of birth $R_{\rm gal}$, producing a broader abundance distribution for any given age. However, even though this prescription is much stronger than the estimates of, e.g., \citet{frankel_measuring_2018}, the model cannot reproduce the observed age--abundance trends in [O/H] and [Fe/H].

Second, as in Section \ref{sec:abundance-distributions}, we investigate a model where the infalling gas is enriched to a metallicity $\mathOH=\mathFeH={\rm [X/H]}_{\rm CGM}$ before accreting onto the disk. Column (c) of Figure \ref{fig:abundance-evolution-params} shows results for the case where ${\rm [X/H]}_{\rm CGM}-0.5$, the highest metallicity allowed by the local low-$\alpha$ population. Pre-enriched infall at this level mitigates but does not completely solve the two discrepancies. The dilution effect of the second infall is reduced to the $\sim0.3$-dex level as the gas which replenishes the Galaxy's reservoir is no longer pristine; however, the width of the stellar abundance distribution at any given age is also reduced, because the enriched gas accretion imposes a lower limit on the metallicity of the outermost regions, from which the stars of the low-metallicity tail in the Solar neighborhood are drawn. The late-time gas abundance evolution is similar to the fiducial model, but it ends at slightly super-Solar metallicity --- an effect which can be compensated by a slightly increased value of $\eta$.

Finally, we explore a model where the local thick-to-thin disk surface density ratio is $\sim4$ times larger, $f_\Sigma(R_\odot)=0.48$. This is higher than most of the constraints from population counts or GCE models (see Section \ref{sec:sfh}). Column (d) of Figure \ref{fig:abundance-evolution-params} shows that requiring a more massive thick disk can reduce the dilution and recent evolution of the ISM, similar to the pre-enriched infall, because more of the gas disk is built up during the first infall phase. The model produces the best agreement with the observed age--[Fe/H] relation (second row). However, agreement with the observed age--[O/Fe] relation is poor, as the [O/Fe] of the old low-$\alpha$ disk is $\sim0.1$ dex lower than observed.

Overall, no modification to the \yZ{1} model is able to completely overcome both the dilution and late-time evolution issues.
